---
title: "Single T-cell Analysis by Rna-seq and Tcr TRACking (Startrac)"
package: Startrac
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc_float: true    
  BiocStyle::pdf_document: default
abstract: |
  Instructions on how to use Startrac
vignette: |
  %\VignetteIndexEntry{Single T-cell Analysis by Rna-seq and Tcr TRACking (Startrac)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction
To be added...

# Quick start
We load the Startrac package and other packages for visualization.
```{r results='hide', message=FALSE, warning=FALSE}
library("Startrac")
library("tictoc")
library("ggpubr")
library("ggplot2")
library("ComplexHeatmap")
library("RColorBrewer")
library("circlize")
```

Then we read in the example data. The input to Startrac should be a data.frame for which the columns 'clone.id', 'patient', 'majorCluster' and 'loc' are required.
```{r}
dat.file <- system.file("extdata/example.cloneDat.Zhang2018.txt",
                        package = "Startrac")
in.dat <- read.table(dat.file,stringsAsFactors = F,head=T)
head(in.dat)
```

Run the pipeline using below command to obtain the indices of the entire data set and data from each patient.
```{r}
tic("Startrac.run")
out <- Startrac.run(in.dat, proj="CRC", cores=NULL,verbose=F)
toc()
```


# Visualization

## cluster level index of all data
```{r fig.width=5,fig.height=7}
ggbarplot(data.table::melt(subset(out$cluster.data,aid=="CRC"),
                           id.vars=c("aid","majorCluster"),
                           variable.name="index"),
          x="majorCluster",y="value",palette = "npg",fill = "index") +
  facet_wrap(~index,ncol=1,scales = "free_y") +
  theme(axis.text.x=element_text(angle = 60,hjust = 1))
```

## cluster level index by patients.
First, we examine the expansion index.
```{r fig.width=5,fig.height=4,warning=FALSE}
ggboxplot(subset(out$cluster.data,aid!="CRC"),
          x="majorCluster",y="expa",
          color = "steelblue", add = "point", outlier.colour=NULL) +
  theme(axis.text.x=element_text(angle = 60,hjust = 1))
```

Similarly, we examine the migration index and transition index
```{r fig.width=5,fig.height=4}
ggboxplot(subset(out$cluster.data,aid!="CRC"),
          x="majorCluster",y="migr",
          color = "steelblue", add = "point", outlier.colour=NULL) +
  theme(axis.text.x=element_text(angle = 60,hjust = 1))
```

```{r fig.width=5,fig.height=4}
ggboxplot(subset(out$cluster.data,aid!="CRC"),
          x="majorCluster",y="tran",
          color = "steelblue", add = "point", outlier.colour=NULL) +
  theme(axis.text.x=element_text(angle = 60,hjust = 1))
```

## pairwise index of all data

migration index between two locations
```{r fig.width=5,fig.height=7}
ggbarplot(data.table::melt(subset(out$pIndex.migr,aid=="CRC"),
                           id.vars=c("aid","majorCluster"),
                           variable.name="pair"),
          x="majorCluster",y="value",palette = "npg",fill = "pair") +
  facet_wrap(~pair,ncol=1,scales = "free_y") +
  theme(axis.text.x=element_text(angle = 60,hjust = 1))
```

transition index between two major clusters
```{r fig.width=8,fig.height=6}
dat.plot <- as.matrix(subset(out$pIndex.tran,aid=="CRC")[,c(-1,-2)])
rownames(dat.plot) <- subset(out$pIndex.tran,aid=="CRC")[,2]
dat.plot[is.na(dat.plot)] <- 0
col.heat <- colorRamp2(seq(0,0.12,length=15),
                       colorRampPalette(rev(brewer.pal(n=7,name="RdBu")))(15),
                       space = "LAB")
Heatmap(dat.plot,name="pIndex.tran",col = col.heat)
```

# Step by Step

Construct the Startrac object using the example data. The input data and some derived information are stored in slots, for exmaple `cell.data` for the original input data, `clonotype.dist.cluster` for the distribution of clone along major clusters, `clonotype.dist.loc` for the distribution of clone along locations. The parameter `aid` is used to identify the dataset, it can be the project name or patient id.  
```{r}
dat.file <- system.file("extdata/example.cloneDat.Zhang2018.txt",
                        package = "Startrac")
in.dat <- read.table(dat.file,stringsAsFactors = F,head=T)
obj <- new("Startrac",in.dat,aid="CRC")
```

We can examine these information
```{r}
obj@clonotype.dist.cluster[1:3,1:4]
obj@clonotype.dist.loc[1:3,]
```

Calculate the cluster level index. The cluster level indices are stored in slot `cluster.data`.
```{r }
obj <- calIndex(obj)
head(obj@cluster.data)
```

Calculate the pairwise index. 
```{r }
tic("pIndex")
obj <- pIndex(obj)
toc()
```

These indices are stored in slot `pIndex.migr` and `pIndex.tran`.
```{r }
head(obj@pIndex.migr)
obj@pIndex.tran[1:3,1:6]
```

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
